---
title: "Cluster Analysis Language And Human Similarity Data"
author: Martin Zettersten 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: yes
---

```{r setup, include = F}

# load packages
library(rmarkdown)
library(knitr)
library(tidyverse)
library(here)
library(Matrix)
library(dendextend)
library(ggdendro)
library(clues)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)
```

## Pre-processing functions

I will make these a bit prettier, but basically they just convert the existing data frames with distance/ similarity info into "distance" matrices.

``` {r}
#### useful functions ####
convert_similarity_to_matrix <- function(wide_data) {
  #work into symmetric matrix (messy)
  #convert to matrix
  m <- as.matrix(wide_data)
  #add extra row
  m <- rbind(m,c(rep(NA,ncol(m)-1),1.0))
  row.names(m)[nrow(m)] <- colnames(m)[ncol(m)]
  #add extra column
  m <- cbind(c(1.0,rep(NA,nrow(m)-1)),m)
  colnames(m)[1] <- row.names(m)[1]
  #convert into symmetric matrix (using forceSymmetric in Matrix package)
  diag(m) <- 1
  m <- forceSymmetric(m)
  #return
  m
}

convert_similarity_to_distance <- function(wide_data, col_name,reverse_dist=T, human_data=T) {
  #convert human similarity values based on a similarity column name
  #extract subset of wide human data
  temp <- as.data.frame(wide_data) %>%
    select(animal1,animal2, !!col_name) %>%
    spread(animal2,!!col_name,fill="", convert=T)
  #change animal1 column to row name
  row.names(temp) <- temp$animal1
  temp <- temp %>%
    select(-animal1)
  if (human_data) {
    #convert to symmetric matrix
    temp <- convert_similarity_to_matrix(temp)
  } else {
    #convert to matrix
    temp <- as.matrix(temp)
  }
  
  if (reverse_dist) {
    #convert from similarity to "distance"
    temp <- 1- temp
  }
  #return
  temp
}
```

##LOAD DATA

```{r}
LANG_ANIMAL_DISTANCE_COLOR <- here("data/processed/animal_color_distances_language_wiki.csv")
LANG_ANIMAL_DISTANCE_COLOR
LANG_ANIMAL_DISTANCE_SHAPE<- here("data/processed/animal_shape_distances_language_wiki.csv")
LANG_ANIMAL_DISTANCE_TEXTURE <- here("data/processed/animal_texture_distances_language_wiki.csv")
TIDY_HUMAN_PATH <- here("data/processed/tidy_human_data.csv")

language_data <- read_csv(LANG_ANIMAL_DISTANCE_COLOR) %>%
  left_join(read_csv(LANG_ANIMAL_DISTANCE_SHAPE), by  = c("animal1", "animal2")) %>%
  left_join(read_csv(LANG_ANIMAL_DISTANCE_TEXTURE),by  = c("animal1", "animal2")) %>%
  select(-contains("PCA"))
human_data <- read_csv(TIDY_HUMAN_PATH)
human_data_wide <- human_data %>%
  unite("measure", c("participant_type", "similarity_type")) %>%
  spread(measure, human_similarity)
```

##COLOR

###SIMPLE DENDOGRAMS

####SIGHTED HUMAN DENDOGRAM

```{r}
sighted_human_color_clust <- human_data_wide %>%
  convert_similarity_to_distance("sighted_human_similarity_color") %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(sighted_human_color_clust, rotate = T) +
  ggtitle("Sighted Human Similarity Color")
```

####BLIND HUMAN DENDOGRAM

```{r}
blind_human_color_clust <- human_data_wide %>%
  convert_similarity_to_distance("blind_human_similarity_color") %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(blind_human_color_clust, rotate = T) +
  ggtitle("Blind Human Similarity Color")
```

####LANGUAGE DENDOGRAM

```{r}
language_color_clust <- language_data %>%
  convert_similarity_to_distance("language_similarity_simple_dist_color", reverse_dist=F, human_data=F) %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(language_color_clust, rotate = T) +
  ggtitle("Language Distances Color")
```

###BLIND TO SIGHTED HUMAN COMPARISON

####DEFAULT TANGLEGRAM

Lower entanglement is better (0 is perfect alignment).

```{r}
dends <- dendlist(as.dendrogram(blind_human_color_clust), as.dendrogram(sighted_human_color_clust))
x <- tanglegram(dends, common_subtrees_color_branches = TRUE, highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(blind_human_color_clust), as.dendrogram(sighted_human_color_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```


####Bk plot

See https://cran.r-project.org/web/packages/dendextend/vignettes/introduction.html#the-fowlkes-mallows-index-and-the-bk-plot for details.

The FM-Index is a measure of similarity between two clusterings (higher means greater similarity). The Bk plot shows the FM-INdex for various values of k, where k is the number of clusters. The dashed black line shows the *expected value assuming no connection between the clusterings*, the red line shows the *critical significance level*. **Dotted points above indicate significantly similar clusterings, i.e. better than one would expect from simply reshuffling labels**.

```{r}
Bk_plot(as.dendrogram(blind_human_color_clust), as.dendrogram(sighted_human_color_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a smaller cluster number (k=5).

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(blind_human_color_clust, k=5), cutree(sighted_human_color_clust, k=5))
```

###SIGHTED TO LANGUAGE COMPARISON

####DEFAULT TANGLEGRAM

Lower entanglement is better (0 is perfect alignment).

```{r}
dends <- dendlist(as.dendrogram(sighted_human_color_clust), as.dendrogram(language_color_clust))
x <- tanglegram(dends, common_subtrees_color_branches = TRUE, highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(sighted_human_color_clust), as.dendrogram(language_color_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a smaller cluster number.

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(sighted_human_color_clust, k=5), cutree(language_color_clust, k=5))
```


###BLIND TO LANGUAGE COMPARISON

####DEFAULT TANGLEGRAM

Lower entanglement is better (0 is perfect alignment).

```{r}
dends <- dendlist(as.dendrogram(blind_human_color_clust), as.dendrogram(language_color_clust))
x <- tanglegram(dends, common_subtrees_color_branches = TRUE, highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(blind_human_color_clust), as.dendrogram(language_color_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a smaller cluster number (k=5).

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(blind_human_color_clust, k=5), cutree(language_color_clust, k=5))
```

##SHAPE

###SIMPLE DENDOGRAMS

####SIGHTED HUMAN DENDOGRAM

```{r}
sighted_human_shape_clust <- human_data_wide %>%
  convert_similarity_to_distance("sighted_human_similarity_shape") %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(sighted_human_shape_clust, rotate = T) +
  ggtitle("Sighted Human Similarity shape")
```

####BLIND HUMAN DENDOGRAM

```{r}
blind_human_shape_clust <- human_data_wide %>%
  convert_similarity_to_distance("blind_human_similarity_shape") %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(blind_human_shape_clust, rotate = T) +
  ggtitle("Blind Human Similarity shape")
```

####LANGUAGE DENDOGRAM

```{r}
language_shape_clust <- language_data %>%
  convert_similarity_to_distance("language_similarity_simple_dist_shape", reverse_dist=F, human_data=F) %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(language_shape_clust, rotate = T) +
  ggtitle("Language Distances shape")
```

###BLIND TO SIGHTED HUMAN COMPARISON

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(blind_human_shape_clust), as.dendrogram(sighted_human_shape_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(blind_human_shape_clust), as.dendrogram(sighted_human_shape_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a small cluster number (k=5).

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(blind_human_shape_clust, k=5), cutree(sighted_human_shape_clust, k=5))
```

###SIGHTED TO LANGUAGE COMPARISON

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(sighted_human_shape_clust), as.dendrogram(language_shape_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(sighted_human_shape_clust), as.dendrogram(language_shape_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a small cluster number (k=5).

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(sighted_human_shape_clust, k=5), cutree(language_shape_clust, k=5))
```


###BLIND TO LANGUAGE COMPARISON

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(blind_human_shape_clust), as.dendrogram(language_shape_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(blind_human_shape_clust), as.dendrogram(language_shape_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a small cluster number (k=5)
```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(blind_human_shape_clust, k=5), cutree(language_shape_clust, k=5))
```

##TEXTURE

###SIMPLE DENDOGRAMS

####SIGHTED HUMAN DENDOGRAM

```{r}
sighted_human_skin_clust <- human_data_wide %>%
  convert_similarity_to_distance("sighted_human_similarity_skin") %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(sighted_human_skin_clust, rotate = T) +
  ggtitle("Sighted Human Similarity skin")
```

####BLIND HUMAN DENDOGRAM

```{r}
blind_human_skin_clust <- human_data_wide %>%
  convert_similarity_to_distance("blind_human_similarity_skin") %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(blind_human_skin_clust, rotate = T) +
  ggtitle("Blind Human Similarity skin")
```

####LANGUAGE DENDOGRAM

```{r}
language_skin_clust <- language_data %>%
  convert_similarity_to_distance("language_similarity_simple_dist_texture", reverse_dist=F, human_data=F) %>%
  as.dist() %>%
  hclust()
ggdendro::ggdendrogram(language_skin_clust, rotate = T) +
  ggtitle("Language Distances skin")
```

###BLIND TO SIGHTED HUMAN COMPARISON

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(blind_human_skin_clust), as.dendrogram(sighted_human_skin_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(blind_human_skin_clust), as.dendrogram(sighted_human_skin_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a small cluster number (k=5).

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(blind_human_skin_clust, k=5), cutree(sighted_human_skin_clust, k=5))
```

###SIGHTED TO LANGUAGE COMPARISON

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(sighted_human_skin_clust), as.dendrogram(language_skin_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(sighted_human_skin_clust), as.dendrogram(language_skin_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a small cluster number (k=5).

```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(sighted_human_skin_clust, k=5), cutree(language_skin_clust, k=5))
```


###BLIND TO LANGUAGE COMPARISON

####TANGLEGRAM AFTER ROTATING TREES TO FIND MORE COMPARABLE SOLUTION
```{r}
dends <- dendlist(as.dendrogram(blind_human_skin_clust), as.dendrogram(language_skin_clust))
x <- dends %>%
  untangle(method = "step2side") %>% 
  tanglegram( common_subtrees_color_branches = TRUE,highlight_branches_lwd = F) 
paste("entanglement = ",round(entanglement(x), 2))
```

####Bk plot

```{r}
Bk_plot(as.dendrogram(blind_human_skin_clust), as.dendrogram(language_skin_clust), main="FM-Index for different numbers of clusters.\nDots closer to 1 indicate more similar clusterings.\nDots above red line are significant.")
```

####Various other metrics of similarity

Choosing a small cluster number (k=5)
```{r}
#five clusters
print("FIVE CLUSTERS")
adjustedRand(cutree(blind_human_skin_clust, k=5), cutree(language_skin_clust, k=5))
```